---
Title: Results
Date: 2023-09-07 11:03:26
LastEditors: Hwrn hwrn.aou@sjtu.edu.cn
LastEditTime: 2023-09-07 16:12:25
FilePath: /2021_09-MT10kSW/results/Rmd/check_results.1.Rmd
Description: results
output:
  html_document:
    df_print: paged
    theme: cerulean
    highlight: haddock
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: yes
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Work/2021_09-MT10kSW/")
```

```{r config, message=FALSE, warning=FALSE}
source("workflow/utils/RLib.local/R/init.r", chdir = TRUE)
```

```{r load_data, dependson=c("config")}
taxon_color <- load__taxon_color()

Stdb <- load__Stdb()
Wtdb <- load__Wtdb()
genome_taxonomy <- load__genome_taxonomy(Stdb, Wtdb)

genome_rltabd <-
  stringr::str_glue("Wtdb.relative_abundance.tsv") %>%
  file_path$file_path$results() %>%
  as.character() %>%
  get_relative_abundance(genome_taxonomy) %>%
  mutate(
    Taxa_label = sapply(taxon.split(get("Taxonomy"), 1, 7), get_taxon_color)
  )
```

```{r genome_prevalence_across_group, dependson=c("load_data")}
annotation_genomes <-
  get_taxon_group(genome_rltabd, "Genome", "Group") %>%
  {
    rownames(.) <- .$Genome %>% gsub("^(.+)$", "\\1.fa", .)
    .$Genome <- NULL
    .
  }
n_genomes_in_group <-
  annotation_genomes %>%
  apply(2, . %>%
    {
      sum(. != "")
    })
genome_prevalence_across_group <-
  annotation_genomes %>%
  split(
    apply(., 1, . %>% paste(collapse = ""))
  ) %>%
  lapply(function(df) {
    n_genomes_in_this_group <- df[1, ] %>%
      .[. != ""] %>%
      n_genomes_in_group[.]
    c(unique(df),
      Genomes = dim(df)[1],
      Percent = (dim(df)[1] / sum(n_genomes_in_this_group)) %>% round(4)
    )
  }) %>%
  bind_rows(.id = "Group_label") %>%
  column_to_rownames("Group_label")

genome_prevalence_across_group
```

## Metagenomic profile
```{r}
sample.log <- "
Site   | Layer | Raw reads pairs | Clean reads pairs
TY.044 | water |     52876904    |     50400486
TY.041 | water |     60268395    |     59411501
TY.040 | water |     54401230    |     52097922
YW.021 | water |     55766716    |     55319739
WQ.022 | water |     68320893    |     67839251
YW.020 | water |     59236043    |     58804951
WQ.024 | water |     76446504    |     75518696
YW.019 | water |     62709177    |     62186093
WQ.021 | water |     65934699    |     65314107
YW.023 | water |     50321128    |     49906256
" %>%
  textConnection() %>%
  {
    data <- read.csv(., sep = "|")
    close(.)
    data
  }

sample.log2 <- "
Site   | SeqNumbers | MaxLength | GenomeSize |  N50
TY.044 |   480653   |   189618  |  530453602 | 1266
TY.041 |   671837   |   272833  |  814160179 | 1289
TY.040 |   669463   |   138538  |  700544315 | 1358
T1L6   |   698160   |    92345  |  625995281 |  830
T1B10  |   963540   |   342441  |  960386196 |  830
T1B11  |   214163   |   345318  |  231596345 |  925
YW.021 |   293147   |  1065244  |  552083548 | 3224
WQ.022 |   192243   |  1592666  |  336384365 | 2437
YW.020 |   308094   |   881722  |  559512074 | 2773
WQ.024 |   309611   |   669602  |  556555548 | 2931
YW.019 |   219702   |   691672  |  368471939 | 2141
WQ.021 |   243095   |  1231312  |  466699531 | 3225
YW.023 |   223371   |   571787  |  440850287 | 3593
T3L11  |  1894425   |   133816  | 1768025214 |  768
T3L8   |   697406   |   110004  |  642045861 |  747
T3L14  |  1189557   |   123418  | 1155399924 |  935
T1L10  |  1894075   |   238282  | 1866414119 |  885
" %>%
  textConnection() %>%
  {
    data <- read.csv(., sep = "|")
    close(.)
    data
  }

sample.log3 <- "
Sample | Layer  | Mapping rate
TY.044 | water  | 27.08
TY.041 | water  | 51.72
TY.040 | water  | 35.41
T1L6   | R1.0-3 | 45.15
T1L6   | R2.0-3 | 41.52
T1B10  | 0-2    | 43.23
T1B10  | 36-38  | 48.71
T1B10  | 44-46  | 52.72
T1B11  | 0-3    | 50.42
YW.021 | water  | 87.98
WQ.022 | water  | 88.35
YW.020 | water  | 87.60
WQ.024 | water  | 88.69
YW.019 | water  | 91.48
WQ.021 | water  | 84.53
YW.023 | water  | 87.60
T3L8   | 0-3    | 41.49
T3L8   | 6-9    | 39.06
T3L8   | 12-15  | 42.91
T3L8   | 18-21  | 27.29
T3L14  | 0-2    | 49.02
T3L14  | 4-6    | 45.96
T3L14  | 6-8    | 49.93
T3L14  | 12-14  | 45.04
T3L14  | 18-20  | 38.65
T3L11  | 0-3    | 49.48
T3L11  | 6-9    | 49.27
T3L11  | 12-15  | 46.87
T3L11  | 18-21  | 35.22
T1L10  | 0-3    | 48.62
T1L10  | 6-9    | 51.58
T1L10  | 12-15  | 48.00
T1L10  | 18-21  | 42.26
" %>%
  textConnection() %>%
  {
    data <- read.csv(., sep = "|")
    close(.)
    data
  }
```

- The shot-gun sequencing of ten near-benthic seawater samples
    yielded `r prettyNum(mean(sample.log$Clean.reads.pairs), big.mark = ",")`
    clean raw read pairs per sample after being trimmed and filtered,
  composing
    `r round(mean(sample.log$Clean.reads.pairs / sample.log$Raw.reads.pairs), 4) * 100`%
    of the raw reads (Methods).
- In addition, we collected 23 metagenomes at seven sediment sampling stations
  (four stations at the bottom and three stations on the slope)
  in the CD for comparison (Fig. 1, Supplemental Table S1).
- All the 33 samples
    were divided into four groups according to sampling site:
      slope seawater (Sw),
      slope sediments (Ss),
      bottom seawater (Bw),
      and bottom sediments (Bs).
- Assembly of the 33 metagenomes above generated
    `r prettyNum(sum(sample.log2$SeqNumbers), big.mark = ",")` scaffolds
    (average length of
      `r round(sum(sample.log2$GenomeSize) / sum(sample.log2$SeqNumbers))`
      bp).

```{r}
Stdb.quality <- with(Stdb, assign_quality(Completeness, Contamination))
```

- Binning of these contigs resulted in `r dim(Stdb)[1]` draft genomes
    with qualities above the middle-quality threshold
      ($\ge 50\%$ completeness and $\le 10\%$ contamination),
    among which `r summary(Stdb.quality)["high"]`
    were above the high-quality threshold
      ($\ge 90\%$ completeness and $\le 5\%$ contamination)
      according to widely-accepted MAG/SAG standards.
- To represent a species-level clade across samples in the study,
  draft genomes were collected from different samples for de-replication,
  and the resulting `r dim(Wtdb)[1]` genomes
    are hereafter referred to as the MAG (metagenome assembled genome).

## Comparison of the prokaryotic community between seawater and sediment on the slope and bottom
```{r}
sample_meta_cross <-
  read.csv("results/reads_diversity/metadata.tsv", sep = "\t", as.is = TRUE) %>%
  mutate(
    X = get("Sample"),
    Layer = ifelse(
      get("Type") == "16s", get("Sample"), gsub("^[^_]+_", "", get("Sample"))
    ),
    Sample = paste0(get("Group"), "_", get("Layer"))
  ) %>%
  merge(unique(sample_meta[c("Location", "Group")])) %>%
  mutate(Group = factor(get("Group"), names(sample_meta_col))) %>%
  .[order(.$Group, .$Sample), ] %>%
  mutate(Group = as.character(get("Group")))

otu_count <-
  "results/reads_diversity/otu.tsv" %>%
  {
    df <- read.csv(., sep = "\t")
    colnames(df) <-
      c("SpeciesID", read.csv(., sep = "\t", header = FALSE)[1, -1])
    df
  } %>%
  pivot_longer(
    !c("SpeciesID"),
    names_to = "X",
    values_to = "ReadsCount"
  ) %>%
  filter(get("ReadsCount") > 0) %>%
  merge(
    "results/reads_diversity/classification.tsv" %>%
      read.csv(sep = "\t") %>%
      mutate(
        Taxonomy = paste(
          get("Domain"), get("Phylum"), get("Class"),
          get("Order"), get("Family"), get("Genus"),
          get("SpeciesID"),
          sep = ";"
        )
      ) %>%
      dplyr::select(c("SpeciesID", "Taxonomy"))
  ) %>%
  merge(sample_meta_cross)

otu_count %>%
  group_by(get("Sample")) %>%
  summarise(Total = sum(get("ReadsCount"))) %>%
  {
    assertthat::assert_that(all(.$Total == 38000))
  } %>%
  invisible()

set.seed(499)
pr1 <-
  genome_rltabd %>%
  filter(get("Group") %in% c("Bw", "Sw")) %>%
  mutate(
    name = taxon.split(get("Taxonomy"), 1, 3),
    Sample1 = paste0(
      ifelse(
        get("Site") %in% c("TY.044", "TY.041"),
        "3μm filtered", "3μm + 0.22μm filtered"
      ),
      "_",
      get("Layer")
    )
  ) %>%
  reshape2::acast(
    formula = formula("name ~ Sample1"),
    value.var = "Relative_abundance",
    fun.aggregate = sum, fill = 0
  ) %>%
  plot.beta.div(
    pname = "relative abundance",
    method = "nmds", dist = "bray", area = "polygon"
  ) +
  labs(title = "NMDS plot of bray distance between water samples\naccording to MAGs abundnace")
pr1
ggsave("results/figs/supp.figx_filtered.svg", pr1, width = 6, height = 4)
```

- The relative abundance of MAGs were used to describe the community structure,
  and `r sum(sample_meta_cross$Type == "16S")` 16S rRNA samples are also used
    for comparison.
  A total of `r length(unique(otu_count$Taxonomy))` ASVs are identified 
    and rarefied to 38000 reads per sample in 16S data.
  Using NMDS to demonstrate
    the differences in species-level taxa among various samples,
  we found that both the 16S and metagenomic data showed
    significant differences between different sample groups,
  except that the Bs and Ss were very similar in the 16S data.
  
- To compare community structure provided by both type of samples,
  we manually corrected the class level annotation
    of metagenomic and 16S data.
  In general, when considering 
    the prevalence of taxa at the class level in the samples,
  NMDS showed clustering patterns that separated samples
    both by environment (water or sediment)
      and sequencing type (16S or metagenome)
      (Binary Jaccard distance, Adonis test, P value < 0.001) (Fig. 2A).
  However, when considering
    the abundance of each class-level taxon in the community
      among the samples of these four groups,
  the Sw samples was much closer to the sediment samples
    but was significantly separated from the Bw samples
    (Bray‒Curtis distance, Adonis test, P value < 0.001) (Fig. 2B).

- With respect to the taxonomic composition
    at the phylum level and at the class level for some specific clades,
  Gammaproteobacteria and Alphaproteobacteria 
    were the most abundant bacterial taxa, 
  and Marinisomatia was also abundant in some samples from the slope.
  Nitrososphaeria 
    (previously reported as Thaumarchaeota,
      a typical ammonia oxidating archaea)
    was the dominant archaea in all samples 
      as well as an abundant taxon in most of the samples
        except for those from the bottom seawater (Fig. 2C, Fig. S2_2?).

```{r}
ref_genome_info <- read.csv("04_bin/03_reference/genome_info.csv")
venn_phylum <-
  genome_rltabd %>%
  mutate(Phylum = get("Taxonomy") %>% taxon.split(1, 2)) %>%
  with(split(get("Phylum"), get("Group"))) %>%
  lapply(unique)
```

- We further constructed a phylogenetic tree by using
    `r sum(genome_prevalence_across_group[(!genome_prevalence_across_group$Ss == "") | (!genome_prevalence_across_group$Sw == ""), "Genomes"])` MAGs
      detected in the slope and
    `r sum(genome_prevalence_across_group[(!genome_prevalence_across_group$Bs == "") | (!genome_prevalence_across_group$Bw == ""), "Genomes"])` in the bottom samples,
  together with `r dim(ref_genome_info)[1]` reference genomes from the GTDB [13] (Methods, Fig. 3).
- These MAGs were distributed into
    `r length(unique(c(venn_phylum$Ss, venn_phylum$Sw)))` phyla in the slope samples and
    `r length(unique(c(venn_phylum$Bs, venn_phylum$Bw)))` phyla in the bottom samples
      (Supplemental Fig. S2, Supplemental Fig. S3).
  The class-level clades across the seawater and sediment at the slope 
    were obviously more numerous and more widely distributed
      than those at the bottom of the CD
        (Fig. 3, Supplemental Fig. S3).
  A detailed comparison of the microbial communities 
    in the seawater and sediment samples 
      from the bottom and the southern slope 
    will be described in the following section.

```{r}
freq.phylum <-
  genome_taxonomy$Taxonomy %>%
  taxon.split(1, 2) %>%
  split(taxon.split(., 1)) %>%
  lapply(as.character) %>%
  lapply(table)
```
- The MAGs encompassed `r length(freq.phylum$Bacteria)` bacterial and `r length(freq.phylum$Archaea)` archaeal phyla (Fig. 1; Table S2).
    The five most abundant bacterial phyla in our data set were
    `r paste(taxon.split(names(head(sort(unlist(freq.phylum), decreasing = TRUE), 4)), 2), collapse = ", ")`, and `r taxon.split(names(sort(unlist(freq.phylum), decreasing = TRUE)[5]), 2)`.
    `r sum(unlist(freq.phylum) == 1)` other phyla were each represented by only a single MAG.
        This included the `r taxon.split(names(unlist(freq.phylum)), 2)[unlist(freq.phylum) == 1]`.

```{r}
v <- list(
  bw.dom =
    genome_rltabd %>%
      filter(get("Group") == "Bw") %>%
      filter(
        grepl(
          paste("Gammaproteobacteria", "Alphaproteobacteria",
            "Actinomycetia", "Bacteroidota",
            sep = "|"
          ),
          get("Taxonomy")
        )
      ) %>%
      group_by(Sample = get("Sample")) %>%
      summarise(Abundance = sum(get("Relative_abundance"))) %>%
      .$Abundance %>%
      summary() %>%
      lapply(. %>% prettyNum(digits = 3)),
  bw.gammaprot =
    genome_rltabd %>%
      filter(get("Group") == "Bw") %>%
      filter(
        grepl(
          paste("Gammaproteobacteria", sep = "|"),
          get("Taxonomy")
        )
      ) %>%
      group_by(Sample = get("Sample")) %>%
      summarise(Abundance = sum(get("Relative_abundance"))) %>%
      .$Abundance %>%
      summary() %>%
      lapply(. %>% prettyNum(digits = 3)),
  s.gammaprot =
    genome_rltabd %>%
      filter(get("Group") %in% c("Bs", "Ss")) %>%
      filter(
        grepl(
          paste("Gammaproteobacteria", sep = "|"),
          get("Taxonomy")
        )
      ) %>%
      group_by(Sample = get("Sample")) %>%
      summarise(Abundance = sum(get("Relative_abundance"))) %>%
      .$Abundance %>%
      summary() %>%
      lapply(. %>% prettyNum(digits = 3)),
  s.dom =
    genome_rltabd %>%
      filter(get("Group") %in% c("Bs", "Ss")) %>%
      filter(
        grepl(
          paste("Nitrososphaeria", "Marinisomatota", "Gemmatimonadota",
            "Planctomycetota", "Chloroflexota",
            sep = "|"
          ),
          get("Taxonomy")
        )
      ) %>%
      group_by(Sample = get("Sample")) %>%
      summarise(Abundance = sum(get("Relative_abundance"))) %>%
      .$Abundance %>%
      summary() %>%
      lapply(. %>% prettyNum(digits = 3)),
  bw.s.dom =
    genome_rltabd %>%
      filter(get("Group") %in% c("Bw")) %>%
      filter(
        grepl(
          paste("Nitrososphaeria", "Marinisomatota", "Gemmatimonadota",
            "Planctomycetota", "Chloroflexota",
            sep = "|"
          ),
          get("Taxonomy")
        )
      ) %>%
      group_by(Sample = get("Sample")) %>%
      summarise(Abundance = sum(get("Relative_abundance"))) %>%
      .$Abundance %>%
      summary() %>%
      lapply(. %>% prettyNum(digits = 3)),
  s.top3 =
    genome_rltabd %>%
      filter(get("Group") %in% c("Sw", "Ss")) %>%
      group_by(Sample = get("Sample"), Class = taxon.split(get("Taxonomy"), 3)) %>%
      summarize(Abundance = sum(get("Relative_abundance"))) %>%
      group_by(Sample = get("Sample")) %>%
      mutate(Rank = rank(-get("Abundance"))) %>%
      filter(get("Rank") <= 3) %>%
      .$Class %>%
      table() %>%
      .[. > 0] %>%
      .[order(-.)] %>%
      head(3),
  s.bw.gammaprot.rate =
    genome_rltabd %>%
      filter(get("Group") %in% c("Sw", "Ss", "Bw")) %>%
      filter(
        grepl(
          paste("Gammaproteobacteria", sep = "|"),
          get("Taxonomy")
        )
      ) %>%
      group_by(Sample = get("Sample"), Group = get("Group")) %>%
      summarise(Abundance = sum(get("Relative_abundance"))) %>%
      group_by(Group = get("Group") == "Bw") %>%
      summarise(Mean = mean(get("Abundance"))) %>%
      column_to_rownames("Group") %>%
      {
        .["FALSE", "Mean"] / .["TRUE", "Mean"]
      }
)
```

- The community composition of the prokaryotes in seawater and sediments were completely distinct in the bottom of the CD (Fig. 1, Supplemental Fig. S2).
  For bottom seawater, Gammaproteobacteria, Alphaproteobacteria, Actinomycetia, and Bacteroidota were dominated
  accounting for more than `r v$bw.dom$Min.`% of the total,
  in which Gammaproteobacteria was the most abundant class-level clade
  with `r v$bw.gammaprot$Min.`-`r v$bw.gammaprot$Max.`% in all seven bottom seawater samples.
  However, the dominant taxa in class level of the bottom sediments are clearly more diverse and in significantly different proportions than those in bottom seawater.
  Although the Gammaproteobacteria is still among the dominant clade,
  its proportion is significant lower with less than `r v$s.gammaprot$Max.`% in the communities of all bottom sediments (Fig. 1C, Supplemental Fig. S2, Supplemental Table S2).
  In contrast, Nitrososphaeria, Marinisomatota, Gemmatimonadota, Planctomycetota (Phycisphaerae and others), Chloroflexota (Dehalococcoidia and others)
  account for a unignorable proportion (up to ~ `r v$s.dom$Max.`%),
  of which the proportions are quite low among all samples in the bottom seawater
  which are less than `r v$bw.s.dom$Max.`%.

```{r, fig.width=8, fig.height=6}
share_col <-
  c("Water only" = "#66E5FF", "Share" = "#91B2BE", "Sediment only" = "#DBD4C5")
assign_share <-
  . %>%
  gsub("^(Bw|Sw)$", "Water only", .) %>%
  gsub("^(Bs|Ss)$", "Sediment only", .) %>%
  unique() %>%
  sort() %>%
  paste(collapse = ", ") %>%
  gsub("^(.+, .+)$", "Share", .) %>%
  factor(names(share_col) %>% rev())
```

```{r, fig.width=16, fig.height=10}
group1_col = c(
  "Bw|Sw",
  "Bw, Sw",
  "Bw, Ss", "Bs, Bw",
  "Bs, Sw", "Ss, Sw",
  "Bs, Bw, Sw", "Bw, Ss, Sw",
  "Bs, Bw, Ss, Sw",
  "Bs, Bw, Ss", "Bs, Ss, Sw",
  "Bs, Ss",
  "Bs|Ss"
) %>%
  {
    col <- c(paletteer::paletteer_d("ggsci::default_igv")[seq_along(.)])
    names(col) <- .
    col
  }

otu_rltabd <-
  otu_count %>%
  group_by(Sample = get("Sample")) %>%
  mutate(Abundance = get("ReadsCount") / sum(get("ReadsCount")) * 100)

otu_rltabd %>%
  group_by(Taxonomy = get("Taxonomy")) %>%
  mutate(
    Group1 = assign_share(get("Group"))
  ) %>%
  group_by(
    Group1 = get("Group1"),
    Layer = get("Layer"),
    Sample = get("Sample"), Group = get("Group")
  ) %>%
  summarise(Abundance = sum(get("Abundance"))) %>%
  as.data.frame() %>%
  get_percent_plot(
    "Abundance", fill.name = "Group1", sample.name = "Layer"
  ) +
  scale_fill_manual(values = share_col) +
  scale_x_discrete(
    limits = filter(sample_meta_cross, get("Type") == "16S")$Layer
  ) +
  theme(
    axis.text.x = element_text(
      color = sample_meta_col[
        filter(sample_meta_cross, get("Type") == "16S")$Group
      ]
    )
  )
  
  otu_rltabd %>% filter(grepl("uncultured", taxon.split(get("Taxonomy"), 6)))
```
